---
title: "Phylogenetic_coverage_RMBL"
author: "Reed"
date: "3/28/2021"
output: pdf_document
---

Although the pie chart has mostly been relegated as a data visualization tool of the past, I think that is occasionally still does have it's uses. A complication with the size o

```{r Install GGTREE if needed, eval = F, echo = F}
install.packages('BiocManager')
BiocManager::install("ggtree") #to download ggtreee use this 
```

```{r library, results = 'hide', message = F, warning = F, echo = FALSE}
library(tidyverse)
library(ggtree)
library(rotl)
library(ggnewscale)
library(taxizedb)
library(ggimage)
```

Molecular import and wrangle. 
```{r import molecular data & wrangle - Molecular}
non_biotic <- read.csv("../data/non_biotic_poll.csv")[,2]

rmbl_plants <- read.csv('../data/gothic_plant_list.csv') %>% 
  group_by(genus) %>% 
  mutate(no_species = n()) %>% 
  dplyr::select(-species) %>% 
  filter(!family %in% non_biotic) %>% 
  ungroup() %>% 
  mutate(taxid = name2taxid(binomial, db = 'ncbi'))

rmbl_genera <- rmbl_plants %>% 
  distinct(genus) %>% 
  pull(genus) 

rm(non_biotic)
```


```{r import and wrangle flora obs, warning = F}

pollen_morphotypes <- read.csv('../data/pollen_morphotype_grp_lookup.csv') %>% 
  mutate(morphotype = str_to_upper(morphotype))

obs <- read.csv("../data/all_interactions.csv") %>% 
  left_join(., pollen_morphotypes, by = c('Plant_spp_morpho' = 'morphotype')) %>% 
  mutate(
    Plant_spp_morpho = ifelse(!is.na(Genus), Genus, Plant_spp_morpho),
    genus = str_split_i(Plant_spp_morpho,  ' ', 1)) %>% 
  select(-Genus)

morpho <- filter(obs, Observation == 'Microscopy')
field <- filter(obs, Observation == 'Field')
molec <- filter(obs, Observation == 'Molecular')

rm(obs, pollen_morphotypes)
```


Retrieve OTT id's to request an Open Tree of Life Phylogeny. We will not be able to retrieve non-monophyletic genera, so we will try to pull out a single species from each genus and search for thata so we can at lest have some thing...

```{r Find Tree, results = 'hide', message = F, warning = F, eval = F}
resolved_names <- tnrs_match_names(rmbl_genera, context = "Land plants")
resolved_names$in_tree <- is_in_tree(resolved_names$ott_id) # this will tell us whether the genus is in the tree. A number of genera will be false, because they are not monophyletic, but we should be able to recover a single species from the genus for our purposes. 

non_monophyletic_genera <- resolved_names %>% 
  filter(in_tree == "FALSE")  %>% 
  pull(unique_name)
non_monophyletic_genera <- as.vector(non_monophyletic_genera)

species_to_q <- rmbl_plants %>% 
  filter(genus %in% non_monophyletic_genera) %>% 
  group_by(genus) %>% 
  sample_n(1) %>% 
  pull(binomial)

rmbl_genera1 <- rmbl_genera[!rmbl_genera %in% non_monophyletic_genera]
rmbl_queries <- c(rmbl_genera1, species_to_q)

rmbl_queries <- rmbl_queries %>% 
  str_replace("Actaea", "actaea rubra") %>%  # I ran the code below, and came back to fix these by hand
  str_replace("Urtica", "urtica dioica") %>%  # fix a few records by hand...
  str_replace("Viola", "Viola praemorsa") %>%  # just threw in random names,
  str_replace("Heuchera", "heuchera cylindrica")

resolved_names <- tnrs_match_names(rmbl_queries, context = "Land plants")
resolved_names <- resolved_names %>% drop_na(ott_id)
resolved_names$in_tree <- is_in_tree(resolved_names$ott_id)
# our tree misses two genera of 298, not bad...
resolved_names <- resolved_names %>% filter(in_tree == "TRUE")

tree <- tol_induced_subtree(ott_ids = resolved_names$ott_id)

tree[["tip.label"]] <- sub("_.*", "",tree[["tip.label"]])
tip_label_order <- as.data.frame(sub("_.*", "",tree[["tip.label"]])) # we need these to annotate the tree. 
colnames(tip_label_order) <- "genus"

ape::write.tree(tree, file = "../data/rmbl_tree", append = FALSE, digits = 10, tree.names = FALSE)

rm(non_monophyletic_genera, rmbl_queries, rmbl_genera1, resolved_names, species_to_q)
```

```{r Plot a phylogenetic tree with named nodes}

tree <- ape::read.tree(file = "../data/rmbl_tree")
tree[["tip.label"]] <- sub("_.*", "",tree[["tip.label"]])
tip_label_order <- as.data.frame(sub("_.*", "",tree[["tip.label"]])) # we need these to annotate the tree. 
colnames(tip_label_order) <- "genus"

rmbl_tree <- ggtree(tree, size=1.0, branch.length='none', layout='circular')

```

```{r remove, echo = F}
rm(non_monophyletic_genera, rmbl_plants, rmbl_genera)
```

Annotate tree and things
```{r prepare field based data, message = F}
field <- field %>% 
  dplyr::select(genus) %>% 
  distinct() %>% 
  mutate(FIELD = 'Field')

grouptreeSEQ <- left_join(tip_label_order, field, by = "genus")%>% 
  mutate(FIELD = replace_na(FIELD, replace = 'Not Observed')) %>% 
  dplyr::select(FIELD)
rownames(grouptreeSEQ) <- tree$tip.label

rm(field)
```

```{r prepare morphological Tree}
morpho_obs <- morpho %>% 
  dplyr::select(genus) %>% 
  distinct() %>% 
  mutate(MICROSCOPY = 'Microscopy')

grouptreeMORPH <- left_join(tip_label_order, morpho_obs, by = "genus")%>% 
  mutate(MICROSCOPY = replace_na(MICROSCOPY, replace = 'Not Observed')) %>% 
  dplyr::select(MICROSCOPY)

rm(morpho, morpho_obs)
```

```{r prepare molecular Tree}

grouptreeOBS <- molec %>% 
  mutate(MOLECULAR = 'Molecular') %>% 
  left_join(tip_label_order, ., by = "genus") %>% 
  mutate(MOLECULAR = replace_na(MOLECULAR, replace = 'Not Observed')) %>% 
  distinct(genus, .keep_all = T) %>% 
  dplyr::select(MOLECULAR)

```


```{r map the Colors to edges}

node_dist <- data.frame(
  node = c(
    c(339:377, 103:151), # Asterales # DONE
    c(449, 230:231), # Alismatales # DONE
    c(432:441, 210:223), # Asparagales # DONE
    c(379:384, 152:159), # Apiales  # DONE
    c(325:329, 88:93), # Boraginales # DONE
    c(271:287, 34:51), # Brassicales # DONE
    c(402:416, 180:195), #Caryophyllales # DONE
    c(245,8:9),  # Celastrales # DONE
    c(179), # Cornales  # DONE
    c(385:389, 160:165),  # Dipsacales # DONE
    c(390:401, 166:178), # Ericales # DONE
    c(259:267, 23:32), # Fabales # DONE
    c(331:337, 95:102), # Gentianales # DONE
    c(54), # geraniales # DONE
    c(307:324, 69:87), # Lamiales # DONE
    c(444:448, 220:229), # Liliales  # DONE
    c(239:244, 1:7), # Malpighiales # DONE
    c(288, 52:53), # Malvales # DONE
    c(289:293, 54:59), # Myrtales # DONE
    c(417:429, 196:209), # Ranunculales # DONE
    c(247:258, 10:22), # Rosales # DONE
    c(33), # sapindales # DONE
    c(294:300, 60:68), # saxifragaceae # DONE
    c(94), # solanales # DONE
    c(210) # zingiberales # DONE
  ),
  order = c(
    rep("Asterales", each = length(c(339:377, 103:151))),
    rep("Alismatales", each = length(c(449, 230:231))),
    rep("Asparagales", each = length(c(432:441, 210:223))),
    rep("Apiales", each = length(c(379:384, 152:159))),
    rep("Boraginales", each = length(c(325:329, 88:93))), # #
    rep("Brassicales", each = length(c(271:287, 34:51))),
    rep("Caryophyllales", each = length(c(402:416, 180:195))), 
    rep("Celastrales", each = length(c(245,8:9))), 
    rep("Cornales", each = length(c(179))),
    rep("Dipsacales", each = length(c(385:389, 160:165))),
    rep("Ericales", each = length(c(390:401, 166:178))),
    rep("Fabales", each = length(c(259:267, 23:32))),
    rep("Gentianales", each = length(c(331:337, 95:102))),
    rep('Geraniales', each = length(c(54))),
    rep("Lamiales", each = length(c(307:324, 69:87))),
    rep("Liliales", each = length(c(444:448, 220:229))),
    rep("Malpighiales", each = length(c(239:244, 1:7))), 
    rep("Malvales", each = length(c(288, 52:53))),
    rep("Myrtales", each = length(c(289:293, 54:59))), 
    rep("Ranunculales", each = length(c(417:429, 196:209))),
    rep("Rosales", each = length(c(247:258, 10:22))),
    rep("Sapindales", each = length(c(33))),
    rep('Saxifragales', each = length(c(294:300, 60:68))),
    rep("Solanales", each = length(c(94))),
    rep('Zingiberales', each = length(c(210)))
  )
)

upper <- data.frame(
  node = c(
    233,
    c(232,423:431),
    234, 
    c(235:236),
    c(237:238, 246),
    c(268:271), 
    c(301:302),
    c(303:304),
    c(305:306, 330),
    c(338,378)),
  order = c(
    'Early',
    rep('Monocots', each = length(c(232,423:431))),
    'Eudicots', 
    rep('Superrosids', each = length(c(235:236))),
    rep('Fabids', each = length(c(237:238, 246))),
    rep('Malvids', each = length(c(268:271))),
    rep('Superasterids', each = length(c(301:302))),
    rep('Asterids', each = length(c(303:304))),
    rep('Lamiids', each = length(c(305:306, 330))),
    rep('Campanulids', each = length(c(338,378)))
  )
)

node_dist <- bind_rows(node_dist, upper)
APalG <- read.csv('../data/APG-hexCodes.csv') 

node_dist <- left_join(node_dist, APalG, by = 'order')

rm(upper, tip_label_order)
```

Finally create the tree here

```{r Create the tree}

labDataDF <- cbind(grouptreeOBS, grouptreeMORPH, grouptreeSEQ)

rm(grouptreeOBS, grouptreeMORPH, grouptreeSEQ)

cols <- c("Field" = "#f652a0",  "Microscopy" = "#4c5270", 
          "Molecular" = "#36eee0", "Not Observed" = 'grey85')

rmbl_tree <- rotate(rmbl_tree, 236)
rmbl_tree <- rotate(rmbl_tree, 237)
rmbl_tree <- rotate(rmbl_tree, 348)
rmbl_tree <- rotate(rmbl_tree, 268)

ob <- gheatmap(rmbl_tree, labDataDF, offset=.8, width=.2,
               colnames_angle=95, colnames_offset_y = .25, colnames = F, color = 'grey85') +
  scale_fill_manual('Status', values = cols ) +

  theme(plot.title = element_text(hjust = 0.5), legend.title.align=0.5, 
        legend.position="bottom") +
  ggtitle("Plant genera \n indicated as interacting with Bombus")

ob1 <- ob %<+% 
  node_dist + 
  aes(color=I(color)) 

ob1 +
  geom_cladelab(node=417, label="Ranunculales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=271, label="Brassicales",  angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=259, label="Fabales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=247, label="Rosales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=390, label="Ericales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=325, label="Boraginales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=379, label="Apiales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) +
  geom_cladelab(node=340, label="Asterales", angle='auto', color='white', fontsize = 3.0,
               offset=7, offset.text = 1, align = T, horizontal = F, hjust = 0.5) 

ggsave('../results/rmbl_interactions_tree.png', device = 'png', width = 5, height = 5, units = 'in')
```


```{r clean up environment}
rm(APalG, ob, ob1, obs, node_dist, labDataDF, rmbl_tree, tree)
```

